Object-based Auditory Scenes Similarity Retrieval and Classification With Wavelet Scattering

### 2016-07-11 paper/structureScene.tex
update section VII A

Studies in cognitive psychology investigating urban sound environment quality pointed out that low level holistic descriptors (perceived or measured) such as global sound level, are not sufficient to fully characterized an acoustic scene \cite{guyot2005urban,kang2006urban}. Instead several papers have reported that some cognitive processes such as sound environment quality perception \cite{dubois2006cognitive} or loudness jugdment \cite{kuwano_memory_2003}  relies upon high level attributes such as the identification and assessment of the sound source which composed the scene. It has been showed that, if available, the complete description of the scene in terms of event occurrences is powerful enough to reliably predict high level cognitive classes, \textit{i.e.} the presence of birds are strong pleasantness indicators and very likely to be heard in parks in urban areas \cite{lafayPartI}. Thus recent researches in sound perception now focus on the study of the contribution of specific sound sources in the assessment of a sound environment \cite{ricciardi2015sound,lavandier2006contribution}, viewing an acoustic scene as ``\,skeleton of events on a bed of texture\,''  \cite{nelken_ear_2013}. \gl{debate about the kind of process underlying perception: holistic / event based perception ?}. \\

From a numerical data processing point of view, a holistic scheme, such as the bag-of-frame approach \cite{aucouturier2007bag}, has a simplicity, but clearly face poor performance on realistic conditions \cite{lagrange:hal-01082501}. Instead the outcomes in sound perception cited above encourage us to adopt an object-based  representation of the scene. \gl{vincent, un petit mot pour motiver le fait que le scattering est un bon outil pour separer les evenements peuplant une scène sonore ?}. \\

In a first attempt to make use of an object-based (\emph{ob}) representation of the scene, we propose the follow a clustering approach. After a features extraction step,  a clustering is performed on each scene features using an enhanced version of the popular K-means algorithm known as K-means$++$ \cite{arthur2007k}, which differs from k-means method in terms of selection of starting points. \\

@@

\begin{itemize}
\item \emph{ob-closest}: the similarity between two scenes is equal to the largest similarity between their centroids.
\item \emph{ob-averaged}: the similarity between two scenes is equal to the average of their centroids similarities.
\item \emph{ob-weighted}: for each scene, each centroid is weighted according to the number of frames belonging to its cluster.\\
\end{itemize}

@@

with $c_N$ the $N$'th nearest neihgbor of the centroid $c_i$. \\

@@

To compute the distance between two signatures $S_u$ and $S_v$, we use a variant version of the Earth Moving Distance ($EMD$), known as $\widehat{EMD}$ and introduced in \cite{pele2008linear}. $EMD$ is a cross-bin histogram distance, widely used in computer vision \cite{zhang2007local}, which compute the minimal cost to be paid to transform one histogram into the other. The advantage of using the $EMD$ is that it does not require bins to be aligned. The alignment problem is solved by using a ``\,ground distance\,'', which is the distance between the bins representatives of the two histograms. In our case, the histograms are the cluster weights $w$, and the bins representative the cluster centroids $c$.\\

@@

with $K^{c(uv)}$ being the slice of $K^{c}$ containing the pairwise similarities between the centroids sets $c^u$ and $c^v$ of the scenes $s_u$ and $s_v$ respectively. As suggested in \cite{pele2009fast}, we set the $\alpha$ parameter to one.\\

To obtain the final similarity measure between the scenes $s_u$ and $s_v$, an extended Gaussian kernels $K^s$ \cite{chapelle1999support,jing2003support} is computed:

@@

Different sets of parameters were tested to compute the MFFCs. We here report only the results for the best setting, which is similar to the one used by the winners of the previous DCASE 2013 challenge \cite{roma2013}. 40 MFCCs (including the first energy coefficient) are computed for windows of 50ms and hops of 25ms, with a limited frequency range going from $27.5$Hz to $1000$Hz. This setting gives us 600 feature vectors per scenes. It has been reported that classifying the full set of frames directly may not be the best approach \cite{rakotomamonjy2015histogram}. In order to reduce the data dimensionality, a pooling step is performed on the MFCCs vectors: each set of MFCCs vectors is dividing into $n$ non-overlapping $t$-seconds long slices. Each $n$ MFCC slices is then averaged over time. As the \emph{ob} approaches involve a clustering step, it is desirable to ensure that each scene is described by circa the same number of vectors,for both the temporal-scattering and the MFCCs, prior to running the clustering. Thus, $t$ is set to 250ms, which gives us 120 MFCCs vectors per scenes.

@@

For the ASS setting, evaluation is performed on the private dataset of the DCASE 2013 challenge. The metric used is the precision at rank $k$ ($p@k$), which is the  average  number  of  items  of  the  same  class among the $k$ closest items averaged for all possible seed items. The $p@k$ is computed for $k=[1,\ldots,9]$. Note that a $p@1$ is equivalent to an accuracy computed based on the labels of the 1st nearest neighbors of the items to classify. For each experimental setting, we report the results for the set of parameters, \emph{i.e.} the number of clusters $m$ for the \emph{ob} approaches and the scaling parameters $N$ of the RBF kernels (see Eq. \ref{eq:kc}) giving the best $p@9$.

>>>>

Studies in cognitive psychology investigating urban sound environment quality pointed out that low level holistic descriptors (perceived or measured) such as global sound level, are not sufficient to fully characterized an acoustic scene \cite{guyot2005urban,kang2006urban}. Instead several papers have reported that some cognitive processes such as sound environment quality perception \cite{dubois2006cognitive} or loudness jugdment \cite{kuwano_memory_2003}  relies upon high level attributes such as the identification and assessment of the sound source which composed the scene. It has been showed that, if available, the complete description of the scene in terms of event occurrences is powerful enough to reliably predict high level cognitive classes, \textit{i.e.} the presence of birds are strong pleasantness indicators and very likely to be heard in parks in urban areas \cite{lafayPartI}. Thus recent researches in sound perception now focus on the study of the contribution of specific sound sources in the assessment of a sound environment \cite{ricciardi2015sound,lavandier2006contribution}, viewing an acoustic scene as ``\,skeleton of events on a bed of texture\,''  \cite{nelken_ear_2013}. \gl{debate about the kind of process underlying perception: holistic / event based perception ?}.

From a numerical data processing point of view, a holistic scheme, such as the bag-of-frame approach \cite{aucouturier2007bag}, has a simplicity, but clearly face poor performance on realistic conditions \cite{lagrange:hal-01082501}. Instead the outcomes in sound perception cited above encourage us to adopt an object-based  representation of the scene. \gl{vincent, un petit mot pour motiver le fait que le scattering est un bon outil pour separer les evenements peuplant une scène sonore ?}.

In a first attempt to make use of an object-based (\emph{ob}) representation of the scene, we propose the follow a clustering approach. After a features extraction step,  a clustering is performed on each scene features using an enhanced version of the popular K-means algorithm known as K-means$++$ \cite{arthur2007k}, which differs from k-means method in terms of selection of starting points.

@@

\begin{itemize}
\item \emph{ob-closest}: the similarity between two scenes is equal to the largest similarity between their centroids.
\item \emph{ob-averaged}: the similarity between two scenes is equal to the average of their centroids similarities.
\item \emph{ob-weighted}: for each scene, each centroid is weighted according to the number of frames belonging to its cluster.
\end{itemize}

@@

with $c_N$ the $N$'th nearest neihgbor of the centroid $c_i$.

@@

To compute the distance between two signatures $S_u$ and $S_v$, we use a variant version of the Earth Moving Distance ($EMD$), known as $\widehat{EMD}$ and introduced in \cite{pele2008linear}. $EMD$ is a cross-bin histogram distance, widely used in computer vision \cite{zhang2007local}, which compute the minimal cost to be paid to transform one histogram into the other. The advantage of using the $EMD$ is that it does not require bins to be aligned. The alignment problem is solved by using a ``\,ground distance\,'', which is the distance between the bins representatives of the two histograms. In our case, the histograms are the cluster weights $w$, and the bins representative the cluster centroids $c$.

@@

with $K^{c(uv)}$ being the slice of $K^{c}$ containing the pairwise similarities between the centroids sets $c^u$ and $c^v$ of the scenes $s_u$ and $s_v$ respectively. As suggested in \cite{pele2009fast}, we set the $\alpha$ parameter to one.

To get the final similarity measure between the scenes $s_u$ and $s_v$, an extended Gaussian kernels $K^s$ \cite{chapelle1999support,jing2003support} is computed:

@@

Different sets of parameters were tested to compute the MFFCs. We here report only the results for the best setting, which is similar to the one used by the winners of the previous DCASE 2013 challenge \cite{roma2013}. 40 MFCCs (including the first energy coefficient) are computed for windows of 50ms and hops of 25ms, with a limited frequency range going from $27.5$Hz to $1000$Hz. This setting gives us 600 feature vectors per scenes. It has been reported that classifying the full set of frames directly may not be the best approach \cite{rakotomamonjy2015histogram}. In order to reduce the data dimensionality, a pooling step is performed on the MFCCs vectors: each set of MFCCs vectors is dividing into $n$ non-overlapping $t$-seconds long slices. Each $n$ MFCC slices is then averaged over time. As the \emph{ob} approaches involve a clustering step, it is desirable to ensure that each scene is described by circa the same number of vectors prior to running the clustering, for both the temporal-scattering and the MFCCs. Thus, $t$ is set to 250ms, which gives us 120 MFCCs vectors per scenes.

@@

For the ASS setting, evaluation is performed on the private dataset of the DCASE 2013 challenge. The metric used is the precision at rank $k$ ($p@k$), which is the  average  number  of  items  of  the  same  class among the $k$ closest items averaged for all possible seed items. The $p@k$ is computed for $k=[1,\ldots,9]$. Note that a $p@1$ is equivalent to an accuracy computed based on the labels of the 1st nearest neighbors of the items to classify. For each experimental setting, we only report the results obtained with the sets of parameters, \emph{i.e.} the number of clusters $m$ for the \emph{ob} approaches and the scaling parameter $N$ of the RBF kernels (see Eq. \ref{eq:kc}) leading to the best $p@9$.
